Supplementary Information

### 2020-11-15 doc/manuscript/MEE_manuscript.Rmd
organizing appendices

---
title: ''
output: 
    bookdown::word_document2: 
      reference_docx: word-styles-reference.docx #make sure this line is indented! see this guide for formatting: https://rmarkdown.rstudio.com/articles_docx.html
bibliography: references.bib
csl: methods-in-ecology-and-evolution.csl

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#Tables
sites = 1
#Figures
schematic=1
SCBI_example=2
clim_sens=3
size=4
full_examples = 5
# Fig5/ table2- DBH/climate interactions
```

@@

We analyzed previously collected tree-ring data from ten sites ranging from 9.15$^\circ$ to 61.30$^\circ$ N latitude and representing a wide range of forest and tree types: tropical broadleaf deciduous and evergreen, temperate broadleaf deciduous and needleleaf evergreen, and boreal needleleaf evergreen (Tables `r sites`, S1-S2).
Nine of these sites (exception: LT) are large forest dynamics plots of the Forest Global Earth Observatory [ForestGEO; @anderson-teixeira_ctfs-forestgeo_2015].
Trees of species represented within the ForestGEO plots were cored within the plot (n=#) and/or in the vicinity (n=#; max distance: # km), following a variety of sampling protocols designed to meet the varied objectives of the original studies (Tables S1, S3). 
In using this diversity of data sources, we ensured that our approach was able to handle challenges presented by varying methodologies and forest types.

@@

**Table 1 | Sites included in this analysis**. Here and throughout, sites are ordered by descending mean annual temperature. Additional site information is provided in Table S1, and species and their sample sizes are detailed in Tables S2-S3. 
![ ](tables_figures/sites_table.png)
```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @vlam_temperature_2014; @maxwell_declining_2016; @sniderhan_growth_2016; @alfaro-sanchez_growth_2017; @helcoski_growing_2019 @gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020-1; @aus_de_ar_tree_2018; @bumann_assessing_2019; @samonil_individual-based_2013; @kaspar_species-specific_nodate; @alexander_potential_2019; @finzi_carbon_2020
---

All tree cores were measured and cross-dated by the original researchers using standard dendrochronological practices. 
From among the full set of *#* original records, we excluded cores for which we detected errors (e.g., labeling inconsistencies, obvious dating errors) that could not be resolved before finalizing the analysis (n=*#*). 
We also excluded records that had to be excluded due to insufficient sample size or anomalous growth patterns, including
(1) species with <7 cores,
(2) cores with <30 years of record,
(3) contiguous portions of cores containing large outliers ($RW$ > mean plus 5 x SD of $RW$ for the entire core),
and (4) the final 20 years prior to death of trees cored dead. 
The final criteria was implemented to avoid periods of growth decline and potentially altered climate sensitivity prior to death [@cailleret_synthesis_2017; @desoto_low_2020].
From analyses including $DBH$ (see below), we further excluded
(1) trees for which we lacked data required to reconstruct $DBH$,
(2) trees for which there was a significant inconsistency between measured $DBH$ and the sum of $RW$'s across the core (*Appendix S1*),
and (3) parts of records where reconstructed $DBH$ (see below) was represented by <3 conspecific trees.
In total, this resulted in inclusion of *#* cores, `r n_cores_dbh` of which could be included in analyses with $DBH$ (Table S3).

For each year in the tree-ring records, we reconstructed $DBH$, as detailed in *Appendix S1*. 
In most cases, when a recent $DBH$ measurement was available, $DBH$ was reconstructed from the outside in. 
In cases where $DBH$ was not available, but when we knew that the core hit pith or could reasonably estimate how far off it was based on the curvature of the rings [@applequist_simple_1958; @duncan_evaluation_1989], $DBH$ was reconstructed from the inside out. 
In either case, we used allometric equations between $DBH$ and bark thickness to account for changes in bark thickness as the tree grew (Appendix S1; Tables S2, S4).

@@

Monthly climate data for 1901-2019 were obtained from CRU v.4.04 [@harris_updated_2014; @harris_version_2020]. 
Variables considered here included average daily minimum, maximum, and mean temperatures ($T_{min}$, $T_{max}$, $T_{mean}$, respectively); potential evapotranspiration ($PET$); precipitation ($PPT$); and precipitation day frequency ($PDF$).
For BCI, we calculated monthly $PPT$ and $PDF$ from daily precipitation readings made on BCI starting in 1929 [@paton_barro_2019].
All ForestGEO climate records used here are archived in the ForestGEO Climate Data Portal, v1.0-alpha [@anderson-teixeira_forestgeoclimate_2020].

@@

Here, for each permutation, `climate` specifies one of the climate drivers in the climate variable group, analyzed over one of all possible combinations of consecutive months over a 15 month period ending near the time of cessation of formation of each annual ring (*Table S1*). (**[ISSUE #51 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/51)**)
*Climwin* runs all potential models to select the best fit (lowest AIC), and does k-fold cross-validation in its computation of AIC to guard against over-fitting [@van_de_pol_identifying_2016]. 
For each group of candidate climate variables, we move forward with the best variable over the time window identified by *climwin* as a candidate climate variable for the multivariate models.

A challenge to this system arose for the sites that have undergone the most rapid changes in climate and tree growth: SC and LT, where trees exhibit significant growth declines attributed to rising temperatures [@sniderhan_growth_2016] and increasing drought (REF), respectively. 
*([ISSUE #25 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/25))*
Problematically, correlating tree growth residuals from which climate-driven trends had been removed against the climate signal with a strong directional trend would not necessarily identify the most relevant climate drivers.
For these sites, we experimented with three approaches to identifying the most important climate drivers (1) the method described above, (2) detrending the climate variables **(AT:prewhitening?)** prior to the climwin step, and (3) excluding *decades with the most pronounced climate change* from the climwin step of the analysis [*sensu*  @sniderhan_growth_2016] (Appendix S3).
After determining that... , here we present results *(method)*.

We verified that this process identified similar climate variable-month combinations as what would be identified using traditional methods for individual species, as detailed in Appendix S2. (**[ISSUE #35 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/35)**)

@@

**Our process picked out similar climate drivers to what would be obtained via traditional methods (Figs. `r SCBI_example`, **S1**; Table S5; Appendix S2), but with the advantage that *climwin* allowed objective selection of the strongest climate drivers and the time windows over which they were most influential.**
The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, each of which was identified by climwin as the top temperature-related driver *at four of the eight sites*. $T_{min}$ was the top driver at only *one site (BCI)*, and $T_{mean}$ was never identified as the top variable within the temperature group (Fig. `r clim_sens`).
Within the precipitation group, precipitation amount ($PPT$) was identified as the top variable most frequently (n=*6 of 9* sites), but it was not uncommon that it was surpassed by precipitation frequency ($PDF$; n=*3 of 8* sites). 
Optimal time windows often coincided with a site's peak growing season (n= *# of 10* for temperature variables, *# of 10* for precipitation variables), but exceptions were common. 
At the *5 lowest latitude* sites (BCI, HKK, LT, CB, and SCBI), the optimal window for precipitation variables spanned $\ge$ 8 months, ending during the peak growing months of the year of ring formation. 
At the *3 highest latitude sites* (HF, ZOF, and SC), the optimal window for precipitation variables was a short ($\le$ 3 months) window during the previous growing season.
Optimal windows for temperature variables tended to be shorter, the longest being a *6*-month period during the summer (wet season) at HKK. 
At two of the higher-latitude temperate sites (HF and Žofín), temperatures were most influential during late winter/ early spring.
There were also a few instances where previous growing season conditions had the strongest influence.

![**Figure 2 | Example comparison of climate sensitivity derived via traditional methods (**a**) and our approach (**b-f**).** Example is for the sensitivity of 14 species at SCBI (codes given in Table S2) to potential evapotranspiration ($PET$). Panel (**a**) shows a matrix of Pearson correlations between ring-width index and monthly climate variables (using the chronologies of Helcoski et al. 2019). Black rectangle represents the period selected by *climwin* as the most influential window. Panels (**b-d**) give statistics for time windows tested in *climwin*, where window open and close indicate months prior to current August, and cells across the lower diaganol indicate single-month tests (akin to panel **a**). Panels (**b**) and (**c**) give values of linear and quadratic terms for each time window, and (**d**) gives the $\Delta AIC$ for each. The time window with the minimum $\Delta AIC$ (0-3 months prior to August, or May-July; black circles), was identified as the optimal window. Panel (**e**) shows the correlation of individual-level residuals to PET, with the function fit in *climwin*. Finally, panel (**f**) shows GLS model output, where $PET$ was a candidate driver variable (along with $PPT$; $DBH$ not included in this model). Plotted are responses of species for which $PET$ was included in the top model, with best-fit polynomials plotted with solid lines when both first- and second-order terms are significant, dashed lines when only one term is significant, and dotted lines when neither is significant. Transparent ribbons indicate 95% confidence intervals.](tables_figures/quilt_comparison.png)

@@

**Interactive effects of $DBH$ and climate were found for #% of species-variable combinations**. 
Interactions were significant for over half of $DBH$-precipitation variable interactions for all three growth metrics (*51% for all*; Table S6).
The majority (#%) of these interactions were positive, indicating that larger trees generally respond more positively (or less negatively) to precipitation or it's frequency (Fig. **6?**).
Interactions between temperature variables and $DBH$ were significant for *~40%* of cases considered (Table S6). 
The majority (#%) of these interactions were positive, indicating that larger trees generally respond less negatively (or more positively) to $T_{max}$ or $PET$ (Fig. **6?**).

@@

>>>>

---
title: ''
output: 
    #pdf_document: 
    bookdown::word_document2: 
      reference_docx: word-styles-reference.docx #make sure this line is indented! see this guide for formatting: https://rmarkdown.rstudio.com/articles_docx.html
bibliography: references.bib
csl: methods-in-ecology-and-evolution.csl

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#Tables
sites = 1
#Figures
schematic=1
SCBI_example=2
clim_sens=3
size=4
full_examples = 5
# Fig5/ table2- DBH/climate interactions

# SUPPLEMENTARY INFO (this chunk must be repeated in MEE_manuscript_SI.Rmd)
#Appendices

DBH_reconstruction <- "S1"
climate_correction <- "S2"
traditional_comparison <- "S3"
changing_climate <- "S4"

DBH_reconstruction_title <- "Appendix S1. Methods for reconstruction of $DBH$"
climate_correction_title <- "Appendix S2. Methods for climate data evaluation and correction"
traditional_comparison_title <- "Appendix S3.  Methods for comparing climwin results with traditional methods"
changing_climate_title <- "Appendix S4.  Dealing with rapidly changing climate and tree growth"

#SI table numbers
site_details = "S1"
species_table = "S2"
sampling_details_table = "S3"
bark_allometries = "S4"
qualitative_results_comparison = "S5"
climate_interactions_summary = "S6"

#SI table captions
site_details_legend <- "Table S1. Site Details."
species_table_legend <-"Table S2. Species analyzed, their characteristics, and bark allometries applied."
sampling_details_legend <- "Table S3. Sampling details for species by site."
bark_allometries_legend <- "Table S4. Allometric equations for bark thickness."
qualitative_results_comparison_legend <- "Table S5. Qualtiative comparison of results from this study with previous studies employing conventional methods."
climate_interactions_summary_legend <- "Table S6. Frequency of $DBH$-climate interactions across all sites and growth metrics."

#SI figure numbers
traditional_comparison = "S1"

#SI figure captions
traditional_comparison_legend <- "Figure S1. Comparison of our approach with traditional methods of identifying climate signals."

```

@@

We analyzed previously collected tree-ring data from ten sites ranging from 9.15$^\circ$ to 61.30$^\circ$ N latitude and representing a wide range of forest and tree types: tropical broadleaf deciduous and evergreen, temperate broadleaf deciduous and needleleaf evergreen, and boreal needleleaf evergreen (Tables `r sites`, `r site_details`, `r species_table`).
Nine of these sites (exception: LT) are large forest dynamics plots of the Forest Global Earth Observatory [ForestGEO; @anderson-teixeira_ctfs-forestgeo_2015].
Trees of species represented within the ForestGEO plots were cored within the plot (n=#) and/or in the vicinity (n=#; max distance: # km), following a variety of sampling protocols designed to meet the varied objectives of the original studies (Tables `r site_details`, `r sampling_details_table`). 
In using this diversity of data sources, we ensured that our approach was able to handle challenges presented by varying methodologies and forest types.

@@

**Table 1 | Sites included in this analysis**. Here and throughout, sites are ordered by descending mean annual temperature. Additional site information is provided in Table `r site_details`, and species and their sample sizes are detailed in Tables `r species_table`-`r sampling_details_table`. 
![ ](tables_figures/sites_table.png)
```{r eval=FALSE, echo=FALSE}
# refs in this table need to go in the nocite field below
```
---
nocite: | 
  @vlam_temperature_2014; @maxwell_declining_2016; @sniderhan_growth_2016; @alfaro-sanchez_growth_2017; @helcoski_growing_2019 @gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020-1; @aus_de_ar_tree_2018; @bumann_assessing_2019; @samonil_individual-based_2013; @kaspar_species-specific_nodate; @alexander_potential_2019; @finzi_carbon_2020
---

All tree cores were measured and cross-dated by the original researchers using standard dendrochronological practices. 
From among the full set of *#* original records, we excluded cores for which we detected errors (e.g., labeling inconsistencies, obvious dating errors) that could not be resolved before finalizing the analysis (n=*#*). 
We also excluded records that had to be excluded due to insufficient sample size or anomalous growth patterns, including
(1) species with <7 cores,
(2) cores with <30 years of record,
(3) contiguous portions of cores containing large outliers ($RW$ > mean plus 5 x SD of $RW$ for the entire core),
and (4) the final 20 years prior to death of trees cored dead. 
The final criteria was implemented to avoid periods of growth decline and potentially altered climate sensitivity prior to death [@cailleret_synthesis_2017; @desoto_low_2020].
From analyses including $DBH$ (see below), we further excluded
(1) trees for which we lacked data required to reconstruct $DBH$,
(2) trees for which there was a significant inconsistency between measured $DBH$ and the sum of $RW$'s across the core (Appendix `r DBH_reconstruction`),
and (3) parts of records where reconstructed $DBH$ (see below) was represented by <3 conspecific trees.
In total, this resulted in inclusion of *#* cores, `r n_cores_dbh` of which could be included in analyses with $DBH$ (Table `r sampling_details_table`).

For each year in the tree-ring records, we reconstructed $DBH$, as detailed in Appendix `r DBH_reconstruction`. 
In most cases, when a recent $DBH$ measurement was available, $DBH$ was reconstructed from the outside in. 
In cases where $DBH$ was not available, but when we knew that the core hit pith or could reasonably estimate how far off it was based on the curvature of the rings [@applequist_simple_1958; @duncan_evaluation_1989], $DBH$ was reconstructed from the inside out. 
In either case, we used allometric equations between $DBH$ and bark thickness to account for changes in bark thickness as the tree grew (Appendix `r DBH_reconstruction`; Tables `r species_table`, `r bark_allometries`).

@@

Monthly climate data for 1901-2019 were obtained from CRU v.4.04 [@harris_updated_2014; @harris_version_2020], and in some cases corrected based on more local records (Appendix `r climate_correction`). 
Variables considered here included average daily minimum, maximum, and mean temperatures ($T_{min}$, $T_{max}$, $T_{mean}$, respectively); precipitation ($PPT$); and, when deemed reliable (Appendix `r climate_correction`), potential evapotranspiration ($PET$) and precipitation day frequency ($PDF$).
*For BCI, we calculated monthly $PPT$ and $PDF$ from daily precipitation readings made on BCI starting in 1929 [@paton_barro_2019].*
All ForestGEO climate records used here are archived in the ForestGEO Climate Data Portal, *v1.0-alpha* [@anderson-teixeira_forestgeoclimate_2020].

@@

Here, for each permutation, `climate` specifies one of the climate drivers in the climate variable group, analyzed over one of all possible combinations of consecutive months over a 15 month period ending near the time of cessation of formation of each annual ring (Table `r site_details`). (**[ISSUE #51 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/51)**)
*Climwin* runs all potential models to select the best fit (lowest AIC), and does k-fold cross-validation in its computation of AIC to guard against over-fitting [@van_de_pol_identifying_2016]. 
For each group of candidate climate variables, we move forward with the best variable over the time window identified by *climwin* as a candidate climate variable for the multivariate models.

A challenge to this system arose for the sites that have undergone the most rapid changes in climate and tree growth: SC and LT, where trees exhibit significant growth declines attributed to rising temperatures [@sniderhan_growth_2016] and increasing drought (REF), respectively. 
*([ISSUE #25 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/25))*
Problematically, correlating tree growth residuals from which climate-driven trends had been removed against the climate signal with a strong directional trend would not necessarily identify the most relevant climate drivers.
For these sites, we experimented with three approaches to identifying the most important climate drivers (1) the method described above, (2) detrending the climate variables **(AT:prewhitening?)** prior to the climwin step, and (3) excluding *decades with the most pronounced climate change* from the climwin step of the analysis [*sensu*  @sniderhan_growth_2016] (Appendix `r changing_climate`).
After determining that... , here we present results *(method)*.

We verified that this process identified similar climate variable-month combinations as what would be identified using traditional methods for individual species, as detailed in Appendix `r traditional_comparison`. (**[ISSUE #35 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/35)**)

@@

**Our process picked out similar climate drivers to what would be obtained via traditional methods (Figs. `r SCBI_example`, **S2**; Table `r qualitative_results_comparison`; Appendix `r traditional_comparison`), but with the advantage that *climwin* allowed objective selection of the strongest climate drivers and the time windows over which they were most influential.**
The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, each of which was identified by climwin as the top temperature-related driver *at four of the eight sites*. $T_{min}$ was the top driver at only *one site (BCI)*, and $T_{mean}$ was never identified as the top variable within the temperature group (Fig. `r clim_sens`).
Within the precipitation group, precipitation amount ($PPT$) was identified as the top variable most frequently (n=*6 of 9* sites), but it was not uncommon that it was surpassed by precipitation frequency ($PDF$; n=*3 of 8* sites). 
Optimal time windows often coincided with a site's peak growing season (n= *# of 10* for temperature variables, *# of 10* for precipitation variables), but exceptions were common. 
At the *5 lowest latitude* sites (BCI, HKK, LT, CB, and SCBI), the optimal window for precipitation variables spanned $\ge$ 8 months, ending during the peak growing months of the year of ring formation. 
At the *3 highest latitude sites* (HF, ZOF, and SC), the optimal window for precipitation variables was a short ($\le$ 3 months) window during the previous growing season.
Optimal windows for temperature variables tended to be shorter, the longest being a *6*-month period during the summer (wet season) at HKK. 
At two of the higher-latitude temperate sites (HF and Žofín), temperatures were most influential during late winter/ early spring.
There were also a few instances where previous growing season conditions had the strongest influence.

![**Figure 2 | Example comparison of climate sensitivity derived via traditional methods (**a**) and our approach (**b-f**).** Example is for the sensitivity of 14 species at SCBI (codes given in Table `r species_table`) to potential evapotranspiration ($PET$). Panel (**a**) shows a matrix of Pearson correlations between ring-width index and monthly climate variables (using the chronologies of Helcoski et al. 2019). Black rectangle represents the period selected by *climwin* as the most influential window. Panels (**b-d**) give statistics for time windows tested in *climwin*, where window open and close indicate months prior to current August, and cells across the lower diaganol indicate single-month tests (akin to panel **a**). Panels (**b**) and (**c**) give values of linear and quadratic terms for each time window, and (**d**) gives the $\Delta AIC$ for each. The time window with the minimum $\Delta AIC$ (0-3 months prior to August, or May-July; black circles), was identified as the optimal window. Panel (**e**) shows the correlation of individual-level residuals to PET, with the function fit in *climwin*. Finally, panel (**f**) shows GLS model output, where $PET$ was a candidate driver variable (along with $PPT$; $DBH$ not included in this model). Plotted are responses of species for which $PET$ was included in the top model, with best-fit polynomials plotted with solid lines when both first- and second-order terms are significant, dashed lines when only one term is significant, and dotted lines when neither is significant. Transparent ribbons indicate 95% confidence intervals.](tables_figures/quilt_comparison.png)

@@

**Interactive effects of $DBH$ and climate were found for #% of species-variable combinations**. 
Interactions were significant for over half of $DBH$-precipitation variable interactions for all three growth metrics (*51% for all*; Table `r climate_interactions_summary`).
The majority (#%) of these interactions were positive, indicating that larger trees generally respond more positively (or less negatively) to precipitation or it's frequency (Fig. **6?**).
Interactions between temperature variables and $DBH$ were significant for *~40%* of cases considered (Table `r climate_interactions_summary`). 
The majority (#%) of these interactions were positive, indicating that larger trees generally respond less negatively (or more positively) to $T_{max}$ or $PET$ (Fig. **6?**).

@@

# Supplementary files

`r DBH_reconstruction_title`

`r climate_correction_title`

`r traditional_comparison_title`

`r changing_climate_title`

`r site_details_legend`

`r species_table_legend`

`r sampling_details_legend`

`r bark_allometries_legend`

`r qualitative_results_comparison_legend`

`r climate_interactions_summary_legend`

### 2020-11-15 doc/manuscript/MEE_manuscript_SI.Rmd
organizing appendices

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#READING IN TABLES TO GET STATS FOR TEXT
path_to_dir <- dirname(dirname(getwd())) # get path to repo (ForestGEO-climate-sensitivity)
```

@@

@@

## Appendix S2.  Methods for comparing climwin results with traditional methods
*([ISSUE #35 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/35))*

@@

## Appendix S3.  Dealing with rapidly changing climate and tree growth

>>>>

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#READING IN TABLES TO GET STATS FOR TEXT
path_to_dir <- dirname(dirname(getwd())) # get path to repo (ForestGEO-climate-sensitivity)

# SUPPLEMENTARY INFO (this chunk must be repeated in MEE_manuscript_SI.Rmd)
#Appendices

DBH_reconstruction <- "S1"
climate_correction <- "S2"
traditional_comparison <- "S3"
changing_climate <- "S4"

DBH_reconstruction_title <- "Appendix S1. Methods for reconstruction of $DBH$"
climate_correction_title <- "Appendix S2. Methods for climate data evaluation and correction"
traditional_comparison_title <- "Appendix S3.  Methods for comparing climwin results with traditional methods"
changing_climate_title <- "Appendix S4.  Dealing with rapidly changing climate and tree growth"

#SI table numbers
site_details = "S1"
species_table = "S2"
sampling_details_table = "S3"
bark_allometries = "S4"
qualitative_results_comparison = "S5"
climate_interactions_summary = "S6"

#SI table captions
site_details_legend <- "Table S1. Site Details."
species_table_legend <-"Table S2. Species analyzed, their characteristics, and bark allometries applied."
sampling_details_legend <- "Table S3. Sampling details for species by site."
bark_allometries_legend <- "Table S4. Allometric equations for bark thickness."
qualitative_results_comparison_legend <- "Table S5. Qualtiative comparison of results from this study with previous studies employing conventional methods."
climate_interactions_summary_legend <- "Table S6. Frequency of $DBH$-climate interactions across all sites and growth metrics."

#SI figure numbers
traditional_comparison = "S1"

#SI figure captions
traditional_comparison_legend <- "Figure S1. Comparison of our approach with traditional methods of identifying climate signals."

```

@@

\newpage
## Appendix S2. Methods for climate data evaluation and correction

@@

## Appendix S3.  Methods for comparing climwin results with traditional methods
*([ISSUE #35 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/35))*

@@

## Appendix S4.  Dealing with rapidly changing climate and tree growth
